Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-07 surfaces-write-up/surface-write-up.tex
eliminated todos/moved to front

@@

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ satisfying \todo{Peter: Why
do we include this $u$?  In the later sections, I don't use this notation and
we should standardize.

Aaron: Isn't it necessary to distinguish the functions? For instance, we could consider $f_0$ either as a function in degree 1 or 2 for $D = \frac{1}{2}x_0$, and we need
$u^d$ to distinguish which degree it lives in.
I think we will need to go through and include the $d$ throughout.
Do you think there's a way to avoid including the $d$?
}

@@

\noindent
form a spanning set for $H^0(\bp^m, dD)$ over $k$. Furthermore, functions 
of the form $u^d \cdot \prod_{i=0}^n f_i^{c_i}$ satisfying
\begin{align}
\label{align:pm-basis}
c_i = -\lfloor d\alpha_i \rfloor && \text{ for } && i > m
\end{align}\todo{Peter: Should be for $i> m+1$ right?  Looking below I see an $f_0$, where as above $f_i$ ranges 1 to $n$.  We should choose whether or not 0 is included in the indexing and fix the section to match that convention.

Aaron: I tried to fix it at $0$ everywhere. If probably missed a few places, so keep your eyes peeled.
}

@@

%\todo{Peter: I don't like having a remark inside of a proof.  Also, it feels
%like this could reasonably be replaced with a better constructed/organized
%proof, so that $e_i$'s are just defined this way initially; let me know if you
%see a good way to do so...  As for Aaron's comment above, we could do something
%like that, except we already defined $e_i$'s to be something different, which
%might be equally/more confusing.
%Aaron: I also don't like this here, just because it isn't really a remark,
%its more of a notational convention, and it doesn't really belong in the proof.
%I also think it's confusing to switch notation in the middle of a proof.
%Earlier in the proof, we should just call the $e_{ij}$ something like $\hat{e_{ij}}$. Then, we should never define $\epsilon_{ij}$, since it's only used in this
%proof, and should really be $e_{ij}$, as I was explaining in my previous todo.
%Instead of defining $\epsilon_{ij}$, just define $e_{ij}$ which are the
%appropriately scaled $\hat{e_{ij}}$ and everything will be much clearer.
%Robin: agree with Aaron here.}

@@

We claim such a $\beta_i$,
can be expressed as a polynomial $\beta_i(f_1,f_2,f_3,f_4)$ with 
$\beta_i(f_1,f_2,f_3,f_4) = f_i$
as elements of $\bk[\Sigma],$ which is possible as $f_1,f_2,f_3,f_4$ generate $\Sigma,$ and hence generate $\bk[\Sigma]$. Furthermore, these
generate all relations, since they allow us to reduce any monomial $\prod_{j =
1}^n f_j^{r_j}$ to the canonical form where $r_j = - \lfloor d \alpha_i
\rfloor$
\todo{Aaron: Am I misremembering, or did we agree on the phone that you would put $i$ subscripts on all the things that depend on $i$?
If not, weren't you at least going to say somewhere which things depend on $i$?
\\
Peter: I still don't entirely agree that explaining the dependency is
important.  What I mean by this is that the proof fixes $i, \sigma,$ etc. and
proves the desired result.  Proving it for each possible choice of course
proves the general result; however, it doesn't seem important to me whether or
not the intermediate variables are the same in the proof for $i'$ and $\sigma'$
than for $i$ and $\sigma$.  Either way the proof is valid.  This is why I have
been somewhat resistant (sorry if annoyingly so...)to adding them in.

On a side note, in some cases actually writing them in would be cumbersome,
e.g. replacing $s_{j,k}$ with $s_{j,k,\sigma}$ (since there is one choice of
$s_{j,k}$ for every $\sigma$.

If you want, can you tell me exactly which variables would be helpful in
understanding the proof to know their dependencies?

Aaron: Ok, I'll stop bugging you about it. I just thought we
should add it because I didn't understand that it depended
on $i$ when I read it. Maybe we can see what David
has to say about it when he reads it?
Also, I don't see why calling $s_{j,k}$ by $s_{j,k,\sigma}$ is cumbersome,
that seems fine to me.
}

@@

\todo{Aaron: Is one of these terms corresponding to $E$ and the other to $\lambda$? If so, can you say which?\\
Peter: No, one corresponds to $r_j$ and one to the fractional part of $s_j - r_j$.

Aaron: Ok, can you give a 1-2 sentence explanation then, just saying
which which corresponds to which?
}

>>>>

\todo{Make sure to include $u^d$ everywhere necessary}
\todo{Make sure to index starting from 0 in $\bp^m$ section}

@@

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ satisfying

@@

\noindent
form a spanning set for $H^0(\bp^m, dD)$ over $k$. Furthermore, functions 
of the form $u^d \cdot \prod_{i=0}^n f_i^{c_i}$ satisfying
\begin{align}
\label{align:pm-basis}
c_i = -\lfloor d\alpha_i \rfloor && \text{ for } && i > m
\end{align}

@@

%\todo{Peter: I don't like having a remark inside of a proof.  Also, it feels
%like this could reasonably be replaced with a better constructed/organized
%proof, so that $e_i$'s are just defined this way initially; let me know if you
%see a good way to do so...  As for Aaron's comment above, we could do something
%like that, except we already defined $e_i$'s to be something different, which
%might be equally/more confusing.
%Aaron: I also don't like this here, just because it isn't really a remark,
%its more of a notational convention, and it doesn't really belong in the proof.
%I also think it's confusing to switch notation in the middle of a proof.
%Earlier in the proof, we should just call the $e_{ij}$ something like $\hat{e_{ij}}$. Then, we should never define $\epsilon_{ij}$, since it's only used in this
%proof, and should really be $e_{ij}$, as I was explaining in my previous note.
%Instead of defining $\epsilon_{ij}$, just define $e_{ij}$ which are the
%appropriately scaled $\hat{e_{ij}}$ and everything will be much clearer.
%Robin: agree with Aaron here.}

@@

We claim such a $\beta_i$,
can be expressed as a polynomial $\beta_i(f_1,f_2,f_3,f_4)$ with 
$\beta_i(f_1,f_2,f_3,f_4) = f_i$
as elements of $\bk[\Sigma],$ which is possible as $f_1,f_2,f_3,f_4$ generate $\Sigma,$ and hence generate $\bk[\Sigma]$. Furthermore, these
generate all relations, since they allow us to reduce any monomial $\prod_{j =
1}^n f_j^{r_j}$ to the canonical form where $r_j = - \lfloor d \alpha_i
\rfloor$

@@
